Elite individuals, insitutions, and economic growth accounting

### 2024-04-10 presentation/main.tex
vF

\documentclass[10pt]{beamer}
\usepackage{pgf}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
% \usepackage{booktabs}
%Information to be included in the title page:
\title{Elite individuals, insitutions, and economic growth accounting}
\author{James Xu}
\institute{ECON 442, Duke University}
\date{2024}
\usefonttheme[onlymath]{serif}

@@

\begin{frame}{Table of Contents}
    \tableofcontents
\end{frame}
\section{Intro}
\begin{frame}
    \frametitle{Research Question}
    \begin{block}{What is the relationship between elite students, academic institutions, and economic growth?}
        \begin{itemize}
            \item Are there outsized returns to economic growth when there are more academic elites?
            \item How do these indicators interact with GDP per capita? Are some more helpful for growth than others?
        \end{itemize}
    \end{block}
    \begin{exampleblock}{Method}
        Through this paper, I investigate the relationship between the number of top-ranked universities, share of top math students,
        IMO scores, and GDP per capita growth to analyze these questions.
    \end{exampleblock}
\end{frame}

@@

\begin{frame}{Imputation results}{School completion rates}
    \centering
    \includegraphics[width=\textwidth]{../build/xgboost.png}
\end{frame}

\begin{frame}{Imputation results}{EIU Democracy Score}
    \centering
    \includegraphics[width=\textwidth]{../build/eiu_xgboost.png}
\end{frame}

@@

\end{frame}
\subsection{Summary}
\begin{frame}{Summary Statistics}
    \begin{table}
        \caption{Summary Statistics}
        \resizebox{\linewidth}{!} {
            \begin{tabular}{lrrrrr}
                & Count & Mean & Std & Min & Max \\
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
               GDP per capita & 441.00 & 28463.99 & 25255.10 & 543.11 & 149461.79 \\
               GDP per capita growth (bps) & 440.00 & 169.06 & 452.68 & -2292.68 & 3303.05 \\
               EIU Democracy Index & 441.00 & 7.25 & 1.77 & 1.93 & 9.93 \\
               Primary Completion & 441.00 & 90.00 & 10.25 & 51.35 & 101.95 \\
               Lower Sec. Completion & 441.00 & 77.89 & 17.14 & 29.21 & 101.97 \\
               Upper Sec. Completion & 441.00 & 62.20 & 18.48 & 18.44 & 97.40 \\
               Population & 441.00 & 35318600.19 & 59445620.16 & 34000.00 & 333287557.00 \\
               ARWU per million pop ranked & 441.00 & 0.26 & 0.33 & 0.00 & 1.54 \\
               IMO score per log pop & 441.00 & 4.72 & 3.49 & 0.00 & 11.79 \\
               PISA math in global 1\% & 441.00 & 0.94 & 1.46 & 0.00 & 14.64 \\
               PISA math & 441.00 & 461.93 & 56.23 & 315.96 & 574.66 \\
               \end{tabular}
        }
    \end{table}
    \begin{table}
        \tiny
        \caption{Pre-imputation Statistics}
        \begin{tabular}{lrrrrr}
         & Count & Mean & Std & Min & Max \\
         \\[-1.8ex]\hline
         \hline \\[-1.8ex]
        EIU Democracy Index & 323.00 & 7.09 & 1.78 & 1.93 & 9.93 \\
        Primary Completion & 157.00 & 90.29 & 10.31 & 51.35 & 100.00 \\
        Lower Sec. Completion & 179.00 & 76.63 & 20.03 & 29.21 & 99.98 \\
        Upper Sec. Completion & 198.00 & 62.32 & 20.58 & 18.44 & 97.40 \\
        \end{tabular}
        \end{table}
\end{frame}

@@

\section{Empirical Results}
\begin{frame}{Model specification}
    Let elite indicators be: $math99, ARWU, IMO$.
    For the sake of concision, let $E_{i,t}$ be a 1 by 4 matrix defined as:
    \[E_{i,t} = 
    \begin{bmatrix}
        math99_{i, t} & ARWU_{i, t} & IMO_{i, t}
    \end{bmatrix}
    \] for country/region $i$ in year $t$.
    Let the coefficients of $E_{i, t}$ be a 3 by 1 matrix called $\lambda$. For control variables, $I_{i, t}$ is the matrix of variables and $\alpha$ is coefficients.
    \begin{equation}
        Y_{i, t} = \beta_0 + \lambda E_{i, t} + \delta GDPpc \times E_{i,t} + \alpha I_{i, t} + T_t + C_i + \epsilon_{i, t}
    \end{equation}
    where $T,C$ represent time and entity dummies respectively and $Y_{i,t}$ be the GDP per capita growth in basis points for a country/region $i$ and year $t$.

    Different specifications where $T_t, C_i$ are added/removed are also tested. Yearly specification without fixed effects also tested.
\end{frame}

\begin{frame}{Relationships}
    \centering
    \includegraphics[width=\textwidth]{../charts/relationships.png}
\end{frame}
\subsection{Regression tables}
\begin{frame}{PISA panel regression}{Highly dependent on model specification due to rich country bias and limited variation}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{5pt}}lcccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{4}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-5}
                \\[-1.8ex] & \multicolumn{1}{c}{Model 1 (base)} & \multicolumn{1}{c}{Model 2} & \multicolumn{1}{c}{Model 3 (Time FE)} & \multicolumn{1}{c}{Model 4 (Time + Entity FE)}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & & -43.491$^{}$ & -40.291$^{}$ & -126.152$^{*}$ \\
                & & (48.245) & (40.349) & (64.627) \\
                 PISA Math 99 x GDP PC & & 0.000$^{}$ & 0.000$^{}$ & 0.001$^{}$ \\
                & & (0.001) & (0.001) & (0.001) \\
                 IMO score per log population & 6.604$^{}$ & 3.629$^{}$ & 4.581$^{}$ & 20.285$^{}$ \\
                & (8.473) & (10.582) & (8.833) & (21.169) \\
                 IMO score x GDP pc & -0.000$^{}$ & 0.000$^{}$ & -0.000$^{}$ & -0.001$^{*}$ \\
                & (0.000) & (0.000) & (0.000) & (0.000) \\
                 ARWU insitutions & -588.767$^{***}$ & -538.997$^{***}$ & -580.030$^{***}$ & -764.617$^{***}$ \\
                & (135.656) & (178.746) & (149.452) & (250.804) \\
                 ARWU insitutions x GDP PC & 0.009$^{***}$ & 0.008$^{**}$ & 0.009$^{***}$ & 0.011$^{***}$ \\
                & (0.003) & (0.003) & (0.003) & (0.004) \\
                 PISA Math & & 1.150$^{}$ & 0.908$^{}$ & 0.495$^{}$ \\
                & & (0.792) & (0.732) & (1.761) \\
                 GDP per capita & -0.003$^{**}$ & -0.003$^{*}$ & -0.003$^{**}$ & 0.000$^{}$ \\
                & (0.001) & (0.002) & (0.001) & (0.005) \\
                 Primary School Completion Rate & -3.554$^{}$ & -11.911$^{***}$ & -5.387$^{}$ & -4.028$^{}$ \\
                & (3.345) & (4.157) & (3.602) & (4.947) \\
                 Lower Sec. Completion Rate & -1.475$^{}$ & -0.571$^{}$ & -1.291$^{}$ & -0.446$^{}$ \\
                & (2.945) & (3.540) & (2.948) & (4.443) \\
                 Upper Sec. Completion Rate & 5.643$^{**}$ & 6.665$^{**}$ & 5.385$^{**}$ & 3.364$^{}$ \\
                & (2.343) & (2.835) & (2.354) & (3.692) \\
                 Democracy Rating & -1.606$^{}$ & 2.270$^{}$ & -7.429$^{}$ & 36.204$^{}$ \\
                & (13.894) & (17.254) & (14.431) & (61.854) \\
                 Time Effects & Yes & No & Yes & Yes \\
                 Fixed Effects & No & No & No & Yes \\
                 Entities & 89 & 89 & 89 & 89 \\
                \hline \\[-1.8ex]
                 Observations & 440 & 440 & 440 & 440 \\
                 $R^2$ & 0.388 & 0.092 & 0.392 & 0.567 \\
                 Adjusted $R^2$ & 0.364 & 0.064 & 0.364 & 0.428 \\
                 Residual Std. Error & 360.908 (df=423) & 437.923 (df=426) & 360.976 (df=420) & 342.461 (df=332) \\
                 F Statistic & 16.727$^{***}$ (df=16; 423) & 3.314$^{***}$ (df=13; 426) & 14.230$^{***}$ (df=19; 420) & 4.066$^{***}$ (df=107; 332) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}
\end{frame}

\begin{frame}{PISA yearly regression}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{1pt}}lcccccccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{8}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-9}
                \\[-1.8ex] & \multicolumn{1}{c}{2003} & \multicolumn{1}{c}{2006} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2012} & \multicolumn{1}{c}{2015} & \multicolumn{1}{c}{2018} & \multicolumn{1}{c}{2022} & \multicolumn{1}{c}{Panel FE}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & -36.263$^{}$ & -323.594$^{**}$ & 128.012$^{}$ & -57.807$^{}$ & 466.275$^{**}$ & -13.249$^{}$ & -185.929$^{**}$ & -126.152$^{*}$ \\
                & (92.257) & (138.527) & (134.080) & (93.739) & (198.856) & (80.330) & (75.932) & (64.627) \\
                 PISA Math 99 x GDP PC & -0.000$^{}$ & 0.004$^{}$ & 0.001$^{}$ & 0.001$^{}$ & -0.009$^{**}$ & 0.000$^{}$ & 0.002$^{}$ & 0.001$^{}$ \\
                & (0.002) & (0.003) & (0.003) & (0.002) & (0.004) & (0.001) & (0.001) & (0.001) \\
                 IMO score per log population & 5.787$^{}$ & 13.706$^{}$ & 7.250$^{}$ & 9.955$^{}$ & 27.849$^{}$ & 2.586$^{}$ & -10.574$^{}$ & 20.285$^{}$ \\
                & (24.329) & (28.793) & (25.375) & (20.089) & (37.305) & (10.918) & (18.736) & (21.169) \\
                 IMO score x GDP pc & -0.000$^{}$ & -0.001$^{}$ & 0.000$^{}$ & 0.000$^{}$ & -0.001$^{}$ & -0.000$^{}$ & 0.000$^{}$ & -0.001$^{*}$ \\
                & (0.001) & (0.001) & (0.001) & (0.001) & (0.001) & (0.000) & (0.000) & (0.000) \\
                 ARWU insitutions & -461.941$^{}$ & -374.847$^{}$ & 298.772$^{}$ & -746.052$^{**}$ & -1132.300$^{}$ & -365.119$^{}$ & -1310.810$^{***}$ & -764.617$^{***}$ \\
                & (443.713) & (420.235) & (453.763) & (292.719) & (718.477) & (288.492) & (455.353) & (250.804) \\
                 ARWU insitutions x GDP PC & 0.004$^{}$ & 0.001$^{}$ & -0.004$^{}$ & 0.008$^{}$ & 0.030$^{**}$ & 0.006$^{}$ & 0.013$^{}$ & 0.011$^{***}$ \\
                & (0.011) & (0.009) & (0.009) & (0.005) & (0.014) & (0.004) & (0.008) & (0.004) \\
                 PISA Math & 0.436$^{}$ & 6.935$^{***}$ & -6.860$^{***}$ & 0.324$^{}$ & -1.519$^{}$ & 1.177$^{}$ & 2.807$^{}$ & 0.495$^{}$ \\
                & (1.552) & (1.906) & (2.386) & (1.489) & (2.828) & (1.002) & (2.056) & (1.761) \\
                 GDP per capita & -0.006$^{}$ & -0.005$^{}$ & -0.000$^{}$ & -0.003$^{}$ & 0.002$^{}$ & -0.007$^{***}$ & -0.003$^{}$ & 0.000$^{}$ \\
                & (0.005) & (0.005) & (0.006) & (0.002) & (0.005) & (0.002) & (0.004) & (0.005) \\
                 Primary School Completion Rate & 7.306$^{}$ & -11.680$^{}$ & 4.589$^{}$ & -15.665$^{}$ & -4.820$^{}$ & -8.787$^{}$ & 4.752$^{}$ & -4.028$^{}$ \\
                & (4.789) & (10.409) & (13.405) & (9.658) & (14.437) & (5.579) & (9.745) & (4.947) \\
                 Lower Sec. Completion Rate & 4.652$^{}$ & 7.506$^{}$ & -0.454$^{}$ & 2.430$^{}$ & -1.228$^{}$ & 6.617$^{*}$ & -13.439$^{}$ & -0.446$^{}$ \\
                & (6.867) & (8.157) & (8.302) & (5.785) & (10.676) & (3.951) & (8.396) & (4.443) \\
                 Upper Sec. Completion Rate & 4.916$^{}$ & 3.930$^{}$ & -2.992$^{}$ & 5.781$^{}$ & 4.631$^{}$ & 0.908$^{}$ & 10.924$^{*}$ & 3.364$^{}$ \\
                & (7.172) & (6.850) & (7.450) & (4.375) & (8.029) & (3.174) & (6.081) & (3.692) \\
                 Democracy Rating & -91.681$^{**}$ & -148.435$^{**}$ & 59.739$^{}$ & 36.375$^{}$ & 25.423$^{}$ & 15.183$^{}$ & -5.316$^{}$ & 36.204$^{}$ \\
                & (39.476) & (56.849) & (46.067) & (30.648) & (51.072) & (16.370) & (28.766) & (61.854) \\
                \hline \\[-1.8ex]
                 Observations & 40 & 56 & 69 & 61 & 65 & 73 & 76 & 440 \\
                 $R^2$ & 0.677 & 0.536 & 0.268 & 0.363 & 0.267 & 0.343 & 0.477 & 0.567 \\
                 Adjusted $R^2$ & 0.516 & 0.392 & 0.095 & 0.187 & 0.081 & 0.198 & 0.368 & 0.428 \\
                 Residual Std. Error & 188.826 (df=26) & 363.704 (df=42) & 411.759 (df=55) & 250.551 (df=47) & 455.710 (df=51) & 178.763 (df=59) & 325.047 (df=62) & 342.461 (df=332) \\
                 F Statistic & 4.193$^{***}$ (df=13; 26) & 3.732$^{***}$ (df=13; 42) & 1.547$^{}$ (df=13; 55) & 2.060$^{**}$ (df=13; 47) & 1.432$^{}$ (df=13; 51) & 2.365$^{**}$ (df=13; 59) & 4.353$^{***}$ (df=13; 62) & 4.066$^{***}$ (df=107; 332) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{8}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}

@@

@@

\begin{frame}{Interpretation}
    \begin{block}{IMO Scores}
        \begin{itemize}
            \item Mostly indicates positive main relationship with GDP per capita growth, but negative interaction with GDP per capita.
            \item Suggests that lower income countries may see more benefit to what IMO scores are a proxy of (non-causal)
        \end{itemize}
    \end{block}

    \begin{block}{PISA math 99 percentile share}
        \begin{itemize}
            \item Very sensitive to model specification and changes between years
            \item Negative main effect but positive interaction effect
        \end{itemize}
    \end{block}

    \begin{block}{ARWU insitutions per million}
        \begin{itemize}
            \item Consistent negative main effect, but positive interaction effect with GDP per capita
            \item Can interpret as OVB, or that higher income countries may see more use from high-ranking institutions
        \end{itemize}
    \end{block}

@@

>>>>

\documentclass[10pt]{beamer}
\usepackage{pgf}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
% \usepackage{booktabs}
%Information to be included in the title page:
\title{Elite individuals, insitutions, and economic growth accounting}
\author{James Xu}
\institute{ECON 442, Duke University}
\date{April 11, 2024}
\usefonttheme[onlymath]{serif}

@@

\begin{frame}{Table of Contents}
    \tableofcontents
\end{frame}
\section{Intro}
\begin{frame}
    \frametitle{Research Question}
    \begin{block}{What is the relationship between elite students, academic institutions, and economic growth?}
        \begin{itemize}
            \item Are there outsized returns to economic growth when there are more academic elites?
        \end{itemize}
    \end{block}
    \begin{exampleblock}{Method}
        Through this paper, I investigate the relationship between the number of top-ranked universities, share of top math students,
        IMO scores, and GDP per capita growth to analyze these questions.
    \end{exampleblock}
\end{frame}

\begin{frame}{Research Question}{Importance and Relevance}
    \begin{block}{Does elite performance matters in education and human capital?}
        \begin{itemize}
            \item Rough measures of human capital such as school enrollment rate, are often used (e.g. Mankiw-Romer-Weil)
            \item Elite students and top university research may give countries a technology advantage
        \end{itemize}
    \end{block}

    \begin{block}{Implications for economic convergence}
        \begin{itemize}
            \item Richer, larger countries will likely have better research output and universities
            \item Technology and trade conflict between China-US have shown that ``hoarding'' technology may widen gap between nations
        \end{itemize}

    \end{block}
\end{frame}

@@

@@

\end{frame}
\subsection{Summary}

@@

\section{Empirical Results}
\begin{frame}{Model specification}
    \small
    Let elite indicators be: $math99, ARWU, IMO$.
    For the sake of concision, let $E_{i,t}$ be a 1 by 4 matrix defined as:
    \[E_{i,t} = 
    \begin{bmatrix}
        math99_{i, t} & ARWU_{i, t} & IMO_{i, t}
    \end{bmatrix}
    \] for country/region $i$ in year $t$.
    For control variables, $I_{i, t}$ is the matrix of variables and $\alpha$ is coefficients.
    \begin{equation}
        Y_{i, t} = \beta_0 + \lambda E_{i, t} + \delta GDPpc \times E_{i,t} + \alpha I_{i, t} + T_t + C_i + \epsilon_{i, t}
    \end{equation}
    where $T,C$ represent time and entity dummies respectively and $Y_{i,t}$ be the GDP per capita growth in basis points for a country/region $i$ and year $t$.
    \textbf{Controls: school completion rates, GDP per capita, PISA average math score, democracy index}
\end{frame}

\begin{frame}{Relationships}
    \centering
    \includegraphics[width=\textwidth]{../charts/relationships.png}
\end{frame}
\subsection{Regression tables}
\begin{frame}{PISA panel regression}{Highly dependent on model specification due to rich country bias and limited variation}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{5pt}}lcccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{4}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-5}
                \\[-1.8ex] & \multicolumn{1}{c}{Model 1 (base)} & \multicolumn{1}{c}{Model 2} & \multicolumn{1}{c}{Model 3 (Time FE)} & \multicolumn{1}{c}{Model 4 (Time + Entity FE)}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & -31.228$^{*}$ & -38.928$^{*}$ & -34.569$^{*}$ & -115.492$^{***}$ \\
                & (16.314) & (20.802) & (17.850) & (36.599) \\
                 IMO score per log population & 10.674$^{*}$ & 4.924$^{}$ & 3.122$^{}$ & -11.158$^{}$ \\
                & (6.363) & (7.470) & (6.441) & (15.728) \\
                 ARWU insitutions & -166.641$^{**}$ & -162.705$^{*}$ & -176.049$^{**}$ & -159.960$^{}$ \\
                & (68.800) & (86.871) & (72.281) & (112.884) \\
                 Time Effects & No & No & Yes & Yes \\
                 Fixed Effects & No & No & No & Yes \\
                 Controls & No & Yes & Yes & Yes \\
                 Entities & 89 & 89 & 89 & 89 \\
                \hline \\[-1.8ex]
                 Observations & 440 & 440 & 440 & 440 \\
                 $R^2$ & 0.037 & 0.077 & 0.374 & 0.553 \\
                 Adjusted $R^2$ & 0.031 & 0.056 & 0.351 & 0.414 \\
                 Residual Std. Error & 445.700 (df=436) & 439.888 (df=429) & 364.790 (df=423) & 346.511 (df=335) \\
                 F Statistic & 5.616$^{***}$ (df=3; 436) & 3.589$^{***}$ (df=10; 429) & 15.813$^{***}$ (df=16; 423) & 3.983$^{***}$ (df=104; 335) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}
        \small
        \begin{itemize}
            \item Introducing controls and time effects do not have a major effect on coefficients; elite indicators are not highly correlated with controls
            \item Country/region fixed effects have small effect on ARWU variable, large effects on PISA math 99 and IMO variables
        \end{itemize}
\end{frame}

\begin{frame}{PISA yearly regression}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{1pt}}lcccccccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{8}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-9}
                \\[-1.8ex] & \multicolumn{1}{c}{2003} & \multicolumn{1}{c}{2006} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2012} & \multicolumn{1}{c}{2015} & \multicolumn{1}{c}{2018} & \multicolumn{1}{c}{2022} & \multicolumn{1}{c}{Panel FE}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & -52.490$^{}$ & -193.475$^{**}$ & 154.062$^{**}$ & 16.575$^{}$ & -34.309$^{}$ & 3.769$^{}$ & -69.681$^{**}$ & -115.492$^{***}$ \\
                & (48.524) & (84.559) & (62.680) & (36.365) & (76.140) & (27.106) & (27.702) & (36.599) \\
                 IMO score per log population & 4.478$^{}$ & -9.112$^{}$ & 14.038$^{}$ & 7.368$^{}$ & -16.622$^{}$ & 2.193$^{}$ & -0.174$^{}$ & -11.158$^{}$ \\
                & (17.327) & (20.741) & (17.012) & (13.983) & (27.567) & (7.722) & (13.474) & (15.728) \\
                 ARWU insitutions & -300.038$^{**}$ & -454.401$^{***}$ & 136.623$^{}$ & -344.990$^{**}$ & 289.270$^{}$ & 61.220$^{}$ & -829.105$^{***}$ & -159.960$^{}$ \\
                & (111.479) & (167.497) & (188.401) & (133.886) & (270.034) & (125.503) & (206.280) & (112.884) \\
                Controls & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
                \hline \\[-1.8ex]
                 Observations & 40 & 56 & 69 & 61 & 65 & 73 & 76 & 440 \\
                 $R^2$ & 0.675 & 0.499 & 0.264 & 0.302 & 0.100 & 0.294 & 0.417 & 0.553 \\
                 Adjusted $R^2$ & 0.562 & 0.387 & 0.137 & 0.162 & -0.067 & 0.180 & 0.327 & 0.414 \\
                 Residual Std. Error & 179.479 (df=29) & 365.156 (df=45) & 402.134 (df=58) & 254.369 (df=50) & 490.971 (df=54) & 180.728 (df=62) & 335.236 (df=65) & 346.511 (df=335) \\
                 F Statistic & 6.011$^{***}$ (df=10; 29) & 4.479$^{***}$ (df=10; 45) & 2.075$^{**}$ (df=10; 58) & 2.158$^{**}$ (df=10; 50) & 0.598$^{}$ (df=10; 54) & 2.581$^{**}$ (df=10; 62) & 4.649$^{***}$ (df=10; 65) & 3.983$^{***}$ (df=104; 335) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{8}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}
        \small
        \begin{block}{Relationship changes from year to-year}
            \begin{itemize}
                \item 2009 is a significant outlier (coefficients now positive than negative, probably due to recession)
                \item Panel methods may ``average out'' the variations in relationship
                \item High variance in $R^2$ suggest that importance of controls, variables of interest may also not be constant
            \end{itemize}
        \end{block}

@@

\begin{frame}{PISA panel regression}{Without imputation}
    \begin{table}[!htbp] \centering
        \resizebox{\linewidth}{!} {
            \begin{tabular}{@{\extracolsep{5pt}}lcccc}
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
                & \multicolumn{4}{c}{\textit{Dependent variable: GDP Per Capita Growth (bps)}} \
                \cr \cline{2-5}
                \\[-1.8ex] & \multicolumn{1}{c}{Model 1 (base)} & \multicolumn{1}{c}{Model 2} & \multicolumn{1}{c}{Model 3 (Time FE)} & \multicolumn{1}{c}{Model 4 (Time + Entity FE)}  \\
                \hline \\[-1.8ex]
                 PISA Math in global P99 & -31.228$^{*}$ & 5.985$^{}$ & 8.659$^{}$ & 25.421$^{}$ \\
                & (16.314) & (22.468) & (21.493) & (84.291) \\
                 IMO score per log population & 10.674$^{*}$ & 6.824$^{}$ & 12.529$^{}$ & 50.506$^{}$ \\
                & (6.363) & (9.440) & (8.631) & (32.187) \\
                 ARWU insitutions & -166.641$^{**}$ & -190.750$^{}$ & -249.208$^{**}$ & -454.384$^{}$ \\
                & (68.800) & (127.229) & (116.476) & (351.559) \\
                 Time Effects & No & No & Yes & Yes \\
                 Fixed Effects & No & No & No & Yes \\
                 Controls & No & Yes & Yes & Yes \\
                \hline \\[-1.8ex]
                 Observations & 440 & 112 & 112 & 112 \\
                 $R^2$ & 0.037 & 0.204 & 0.383 & 0.715 \\
                 Adjusted $R^2$ & 0.031 & 0.125 & 0.294 & 0.355 \\
                 Residual Std. Error & 445.700 (df=436) & 236.761 (df=101) & 212.750 (df=97) & 203.207 (df=49) \\
                 F Statistic & 5.616$^{***}$ (df=3; 436) & 2.587$^{***}$ (df=10; 101) & 4.294$^{***}$ (df=14; 97) & 1.987$^{***}$ (df=62; 49) \\
                \hline
                \hline \\[-1.8ex]
                \textit{Note:} & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
                \end{tabular}
        }
        \end{table}
        \small
        \begin{itemize}
            \item Relationships are very different from imputed version
            \item Some countries are missing in this regression and some years as well
            \item School completion rates are only available for certain years and we have seen the relationship to be volatile year-to-year
        \end{itemize}
\end{frame}

@@

\begin{frame}{Interpretation}
    \begin{block}{IMO Scores}
        \begin{itemize}
            \item Mostly indicates positive main relationship with GDP per capita growth (large standard errors)
            \item Negative relationship when fixed effects are included, volatile between years
        \end{itemize}
    \end{block}

    \begin{block}{PISA math 99 percentile share}
        \begin{itemize}
            \item Generally negative relationship with GDP per capita growth
            \item Negative relationship when fixed effects are included, volatile between years
        \end{itemize}
    \end{block}

    \begin{block}{ARWU insitutions per million}
        \begin{itemize}
            \item Consistent negative relationship found in regression
            \item To some degree volatile between years, but is most consistent
        \end{itemize}
    \end{block}

@@

\begin{frame}{Appendix: Summary Statistics}
    \begin{table}
        \caption{Summary Statistics}
        \resizebox{\linewidth}{!} {
            \begin{tabular}{lrrrrr}
                & Count & Mean & Std & Min & Max \\
                \\[-1.8ex]\hline
                \hline \\[-1.8ex]
               GDP per capita & 441.00 & 28463.99 & 25255.10 & 543.11 & 149461.79 \\
               GDP per capita growth (bps) & 440.00 & 169.06 & 452.68 & -2292.68 & 3303.05 \\
               EIU Democracy Index & 441.00 & 7.25 & 1.77 & 1.93 & 9.93 \\
               Primary Completion & 441.00 & 90.00 & 10.25 & 51.35 & 101.95 \\
               Lower Sec. Completion & 441.00 & 77.89 & 17.14 & 29.21 & 101.97 \\
               Upper Sec. Completion & 441.00 & 62.20 & 18.48 & 18.44 & 97.40 \\
               Population & 441.00 & 35318600.19 & 59445620.16 & 34000.00 & 333287557.00 \\
               ARWU per million pop ranked & 441.00 & 0.26 & 0.33 & 0.00 & 1.54 \\
               IMO score per log pop & 441.00 & 4.72 & 3.49 & 0.00 & 11.79 \\
               PISA math in global 1\% & 441.00 & 0.94 & 1.46 & 0.00 & 14.64 \\
               PISA math & 441.00 & 461.93 & 56.23 & 315.96 & 574.66 \\
               \end{tabular}
        }
    \end{table}
    \begin{table}
        \tiny
        \caption{Pre-imputation Statistics}
        \begin{tabular}{lrrrrr}
         & Count & Mean & Std & Min & Max \\
         \\[-1.8ex]\hline
         \hline \\[-1.8ex]
        EIU Democracy Index & 323.00 & 7.09 & 1.78 & 1.93 & 9.93 \\
        Primary Completion & 157.00 & 90.29 & 10.31 & 51.35 & 100.00 \\
        Lower Sec. Completion & 179.00 & 76.63 & 20.03 & 29.21 & 99.98 \\
        Upper Sec. Completion & 198.00 & 62.32 & 20.58 & 18.44 & 97.40 \\
        \end{tabular}
        \end{table}
\end{frame}

\begin{frame}{Appendix: Imputation results}{School completion rates}
    \centering
    \includegraphics[width=\textwidth]{../build/xgboost.png}
\end{frame}

\begin{frame}{Appendix: Imputation results}{EIU Democracy Score}
    \centering
    \includegraphics[width=\textwidth]{../build/eiu_xgboost.png}
\end{frame}
